Exploring the Design Space of Differentially Private Bayesian Inference under Hellinger Distance

### 2019-01-20 ICML2019/main.tex

\subsection{A new family of mechanisms}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{??} where the scoring function is based on a metric over
probability distributoins.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\bprior=\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset\equiv\{\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Where $\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)$ denotes the posterior beta distribution inferred starting from the parameters in input.
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

\subsubsection{A first approach: the $\expmech$ mechanism}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot u(x,r)}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS \equiv 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}
Lemma \ref{??} gives us a finite bound on $GS$, while lemma \ref{??} proves privacy $\epsilon$ differential privacy of $\expmech$.

\subsubsection{A second approach based on local sensitivity: $\lexpmech$}
\label{subsec_emls}
Figure \ref{??} show that  the accuracy of $\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively big and hence the variance of the distribution $\expmech$ ends up also being too big. In order to solve this issue we
will proceed in two steps. We will first define $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot u(x,r)}{2\cdot LS(\dataobs)})}
{\displaystyle\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)\equiv\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix \ref{??}.

@@

\subsection{Privacy of Exponential Mechanism Family}
\subsubsection{$\expmech$ $\epsilon-$Differential Privacy}
Standard exponential mechanism is $\epsilon-$differential privacy by \cite{dwork2014algorithmic}.

>>>>

\subsection{A new family of mechanisms}
In this section, we explore  mechanism which we denote by $\expmech$. $\expmech$ is an instance
of the exponential mechanism \cite{??} where the scoring function is based on a metric over
probability distributoins.
The idea is that $\expmech$ will sample and output a distribution with higher probability when
it is closer to the real posterior; importantly,  \emph{closer} is measured using a metric over probability
distributions as opposed to a metric over the parameters that represent it. In this work we will
use the Hellinger distance, but this approach could be extended to other $f$-divergences. We will
start by describing how the mechanism works in the case of the Beta-Binomial model noticing that
the extension to the Dirichelet-Multinomial model is straightforward.
Given a prior distribution $\betad(\alpha, \beta)$ and an observed data set: $\dataobs\in\{0,1\}^n$,
we define the follwing set  where the mechanism will sample from:
\[
  \candidateset\equiv\{\bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta),\dataobs)\mid \dataobs \in \{0,1\}^n\}.
\]
Where $\bysinfer(\binomial{n}{\theta}, \bprior,\dataobs)$ denotes the posterior beta distribution inferred starting from the parameters in input.
Notice that the above set reduces to:
\[
  \{\betad(\alpha',\beta')\mid \alpha'=\alpha + \Delta\alpha, \beta'=\beta + n-\Delta\alpha\}
\]
We use $\bysinfer(\dataobs)$ to denote the Bayesian inference process on the observed data $\dataobs$
when the other two parameters, that is prior and likelihood, are fixed.

% <<<<<<< HEAD
% \subsubsection{Exponential Mechanism with Hellinger Distance Metric ($\expmech$)}
% Our exploration starts from the standard exponential mechanism $\expmech(\dataobs,u,\candidateset)$, which samples an element from the candidate set $\candidateset$ with probability proportional to $\exp(\frac{\epsilon u(\dataobs,r)}{2 GS})$:
% \[
% \underset{z \thicksim \expmech(\dataobs,u,\candidateset) }{\Pr}[z=r] = \frac
% {\exp(\frac{- \epsilon u(\dataobs,r)}{2 GS})}
% {\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon u(\dataobs,r')}{2 GS})},
% \]
% where $u(\dataobs,r)$ is the Hellinger scoring function over candidates, $\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function (i.e. the global sensitivity of the Bayesian inference w.r.t the Hellinger distance), calculated by:
% =======
\subsubsection{A first approach: the $\expmech$ mechanism}
$\expmech(x,u,\candidateset)$  outputs an element $r\in\candidateset$ with  probability: % proportional to $\exp(\frac{\epsilon u(x,r)}{2 GS})$.
\[
%\underset{z \thicksim \expmech(x,u,\candidateset) }{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon\cdot u(x,r)}{2 GS})}
{\Sigma_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot GS})},
\]
where $u(x,r)$ is the Hellinger scoring function: $-\hellinger(\bysinfer(\dataobs), r)$, and $GS$ is the global sensitivity of the scoring function, defined as:
\begin{equation*}
GS \equiv 
\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}}\max_{\{r\in \mathcal{R}\}}
|\hellinger(\bysinfer(\dataobs), r) - \hellinger(\bysinfer(\dataobs'), r)|
\end{equation*}

  \begin{algorithm}
  \caption{$\expmech$ in Beta-binomial model}
  \label{mech:expmech}
  \begin{algorithmic}
  \INPUT observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \STATE {\bf return} $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

Lemma \ref{??} gives us a finite bound on $GS$, while lemma \ref{??} proves privacy $\epsilon$ differential privacy of $\expmech$.

\subsubsection{A second approach based on local sensitivity: $\lexpmech$}
\label{subsec_emls}

Figure \ref{??} show that  the accuracy of $\expmech$ is poor if compared with the one of $\ilapmech$. This is due to the fact that
$GS$ is relatively big and hence the variance of the distribution $\expmech$ ends up also being too big. In order to solve this issue we
will proceed in two steps. We will first define $\lexpmech$, which instead of using the global sensitivity, will use the local sensitivity to calibrate
the noise. This is known to be a non private approach, but it will guide us towards new insights.
We will then proceed to use a smooth bound the local sensitivity to
calibrate the noise.

$\lexpmech$ shares candidate set and utility function with $\expmech$,
and outputs a result $r \in \mathcal{R}$ with probability:
\[
%\underset{z \thicksim \lexpmech(x,u,\candidateset)}{\Pr}[z=r] = \frac
\frac
  {\exp(\frac{\epsilon \cdot u(x,r)}{2\cdot LS(\dataobs)})}
{\displaystyle\sum_{r' \in \mathcal{R}}\exp(\frac{\epsilon\cdot u(x,r')}{2\cdot LS(\dataobs)})},
\]
where $LS(\dataobs)$ is the local sensitivity of the scoring function, defined as:
\begin{equation}\label{eq:local_sensitivity_scoring}
LS(\dataobs)\equiv\max_{\dataobs' \in \datauni^n: \\ \adj{\dataobs}{\dataobs'}, r\in \mathcal{R}}\lvert \hellinger(\bysinfer(\dataobs'), r) - \hellinger(\bysinfer(\dataobs'), r)\rvert 
\end{equation}
The above can be proven to simplify to:
\begin{lem}
\label{lem_loc_opt}
\[
LS(\dataobs) = \max\limits_{\dataobs' \in \datauni^n:\adj{\dataobs}{\dataobs'}} \hellinger(\bysinfer(\dataobs), \bysinfer(\dataobs')).
\]
\end{lem}
The interested reader can check the proof in the Appendix \ref{??}.

@@

\subsection{Privacy of Exponential Mechanism Family}
\subsubsection{$\expmech$ $\epsilon-$Differential Privacy}
\begin{lem}
$\expmech$ is $\epsilon-$differentially private by \cite{dwork2014algorithmic}.
\end{lem}
